---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
```

## Loading and preprocessing the data
```{r, results='markup', warning=TRUE, message=TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
data <- read.csv('activity.csv')
```

## What is mean total number of steps taken per day?
```{r}
nrOfStepsPerDay <- tapply(data$steps, data$date, sum, na.rm=TRUE)
```

### 1. Make a histogram of the total number of steps taken each day
```{r}
qplot(nrOfStepsPerDay, xlab='Total number of steps per day', ylab='Frequency - 500', binwidth=500, fill=I("red"), col=I("blue"))
```

### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
nrOfStepsPerDayMean <- mean(nrOfStepsPerDay)
nrOfStepsPerDayMedian <- median(nrOfStepsPerDay)
```
* Mean number of steps per day: `r nrOfStepsPerDayMean`
* Median number of steps per day:  `r nrOfStepsPerDayMedian`

-----

## What is the average daily activity pattern?
```{r}
avgStepsPerInterval <- aggregate(x=list(stepsMean=data$steps), by=list(interval=data$interval), FUN=mean, na.rm=TRUE)
```

### 1. Make a time series plot
```{r}
ggplot(data=avgStepsPerInterval, aes(x=interval, y=stepsMean)) +
    geom_line(color='red', size=2) +
    xlab("Interval - 5 minute") +
    ylab("Avg nr of steps") 
```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
stepsMax <- which.max(avgStepsPerInterval$stepsMean)
stepsMaxInterval <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", avgStepsPerInterval[stepsMax,'interval'])
```

* Interval with the most steps: `r stepsMaxInterval`

----

## Imputing missing values
### 1. Calculate and report the total number of missing values in the dataset 
```{r}
nrOfMissingValues <- length(which(is.na(data$steps)))
```

* Number of missing values: `r nrOfMissingValues`

### 2. Devise a strategy for filling in all of the missing values in the dataset.

* Using the IMPUTE function to replace missing N/A data with Mean values


### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
dataImputed <- data
dataImputed$steps <- impute(data$steps, fun=mean)
```


### 4. Make a histogram of the total number of steps taken each day 
```{r}
nrOfStepsPerDayImputed <- tapply(dataImputed$steps, dataImputed$date, sum)
qplot(nrOfStepsPerDayImputed, xlab='Total steps per day - Imputed', ylab='Frequency - 500', binwidth=500, fill=I("red"), col=I("blue"))
```

### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
nrOfStepsPerDayMeanImputed <- mean(nrOfStepsPerDayImputed)
nrOfStepsPerDayMedianImputed <- median(nrOfStepsPerDayImputed)
```
* Mean (Imputed): `r nrOfStepsPerDayMeanImputed`
* Median (Imputed):  `r nrOfStepsPerDayMedianImputed`


----

## Are there differences in activity patterns between weekdays and weekends?
### 1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
dataImputed$dateType <-  ifelse(as.POSIXlt(dataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

### 2. Make a panel plot containing a time series plot

```{r}
avgdataImputed <- aggregate(steps ~ interval + dateType, data=dataImputed, mean)
ggplot(avgdataImputed, aes(interval, steps)) + 
    geom_line(color='red', size=2) +
    facet_grid(dateType ~ .) +
    xlab("Interval - 5 minutes") + 
    ylab("Avg nr of steps")
```
